Quantum H theorem, variational entropy, and relaxation processes

### 2021-01-11 tex/quantum.tex
Rewording: finished 1st round of sec. 3.

\color{blue}

%
Now, the summation over $M$ will be expressed in two summations, such as in the
classical case
%
\begin{equation}
    \frac{d\mathcal{H}}{dt}=\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} g^{+}_{nJ}\dot{g}^{+}_{nJ}\de V_J \delta \epsilon_n+\sum^{P}_J  g^{-}_{nJ}\dot{g}^{-}_{nJ} \de V_J\delta \epsilon_n \right), \label{cambioH5}
\end{equation}
%
where $L+P=K$. Besides, $\dot{g}^{+}_{J}$ represents the change on the
deviation on cells that have an excess of particles or energy while
$\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have
missing particles or energy.
On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells
that have an excess of particles or energy. In contrast, $g^{-}_{J}$ represents
the value of the deviation on cells that have missing particles or energy.

Also, on the one hand, $\dot{g}^{+}_{nJ}<0$ describes the loss of particles
and/or energy and so, $g^{+}_{nJ}>0$. On the other hand, $\dot{g}^{-}_{nJ}>0$
describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$.

We sort the previous ideas in the following form
%
\begin{eqnarray}
   &&g^{+}_{nJ}=+|g^{+}_{nJ}|; \ \ \  \dot{g}^{+}_{nJ}=-|\dot{g}^{+}_{nJ}| \nonumber \\
   &&g^{-}_{nJ}=-|g^{-}_{nJ}|; \ \ \ \dot{g}^{-}_{nJ}=+|\dot{g}^{-}_{nJ}| \label{separacion},
\end{eqnarray}
%

and consequently, (\ref{cambioH5}) obtains the following form
%
\begin{equation}
    \frac{d\mathcal{H}}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L} |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\de V_J \delta \epsilon_n+\sum^{P}_J  |g^{-}_{nJ}||\dot{g}^{-}_{nJ}| \de V_J \delta \epsilon_n \right), \label{cambioH6}
\end{equation}
%
and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then
$\frac{d\mathcal{H}}{dt}<0$. With this, we proved that any quantum ideal gas perturbed
(out of equilibrium but not so far from it), evolves such that
$\frac{d\mathcal{H}}{dt}<0$.

When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot
g_{nM}=0$ and finally from (\ref{eq:cambioH1}) $\frac{d\mathcal{H}}{dt}=0$. Then, we can
say that any ideal quantum gas perturbed always evolves to the equilibrium
state, such as the classical $H$-theorem. Then, we obtained the quantum version
of the Boltzmann's $H$-theorem including inhomogeneities in the system.

We obtained some results from the Statistical Mechanic in Quantum Mechanics
scheme. However, the correspondence principle is important to recover all
results from Classical Mechanics. This correspondence principle will be
discussed in the next section.

  %-----------------------------------------

%---------------------------------------------

>>>>

Succinctly, considering a diluted quantum gas contained in a vessel of volume $V$ (divided into
$K$ small cells), with total energy $E$ and $N$ quantum free particles, which initially
is slightly out of equilibrium ---so as to the local equilibrium hypothesis is valid---,
the functional
%
\begin{eqnarray}\label{eq:qHfunctTheo}
    \mathcal{H} (t)&=&\sum_{M=1}^{K} \sum_{n} \bigg[ f_{Mn}(t) \ln f_{Mn}(t)\pm \Big(1 \mp f_{Mn}(t)) \ln (1 \mp f_{Mn}(t)\Big) \Big]   \delta V_M,
\end{eqnarray}
%
where $f_{Mn}$ is the $M$-th cell distribution function,
satisfies $d\mathcal{H}/dt\leq0$, and the equality condition is attained when the system approaches
the equilibrium state at $t\to\infty$. In Eq.~(\ref{eq:qHfunctTheo}), and for a Fermi (Bose)
gas, $f_{Mn}$ corresponds to the Fermi-Dirac (Bose-Einstein) distrubution function for each
cell and locally each cell is in equilibrium, although the complete system may be non-homogeneous,
and is caracterized by the respective $f_{Mn}$, number of particles $\mathcal{N}_M$, energy $\mathcal{E}_M$,
temperature $T_M$, and Legendre Multipliers $\alpha_M$ and $\beta_M$.
